---
title: "Portofolio for Computational Musicology 2025"
author: "Thomas Manu"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    orientation: columns
---

```{r setup, include=FALSE}
library(rjson)
library(tidyverse)
library(tuneR)
library(seewave)
library(ggplot2)
library(dplyr)
library(purrr)
```

# Introduction {.storyboard}

##  {data-width="400"}

### Homework 9

```{r}

# Normalize vectors using various normalization strategies
compmus_normalise <- function(v, method = "euclidean") {
  harmonic <- function(v) v * sum(1 / abs(v))
  manhattan <- function(v) v / sum(abs(v))
  euclidean <- function(v) v / sqrt(sum(v^2))
  chebyshev <- function(v) v / max(abs(v))
  clr <- function(v) { lv <- log(v); lv - mean(lv) }
  softmax <- function(v) exp(v) / sum(exp(v))

  METHODS <- list(
    identity = identity,
    harmonic = harmonic,
    manhattan = manhattan, L1 = manhattan,
    euclidean = euclidean, L2 = euclidean,
    chebyshev = chebyshev, maximum = chebyshev,
    aitchison = clr, clr = clr,
    softmax = softmax
  )

  if (!is.na(i <- pmatch(method, names(METHODS)))) {
    METHODS[[i]](v)
  } else {
    stop("The method name is ambiguous or unsupported.")
  }
}

# Compute chroma features
compmus_chroma <- function(file, norm = "identity") {
  json_data <- fromJSON(file = file)
  
  json_data$tonal$hpcp %>%
    map(~ {
      .x %>%
        circshift(-8) %>%
        matrix(3, 12) %>%
        colMeans() %>%
        compmus_normalise(method = norm) %>%
        enframe(name = "pc") %>%
        mutate(pc = pc - 1)
    }) %>%
    list_rbind(names_to = "time") %>%
    mutate(time = 2048 * time / 44100)
}

# Compute MFCCs (Timbre)
compmus_mfccs <- function(file, norm = "identity") {
  json_data <- fromJSON(file = file)

  json_data$lowlevel$mfcc %>%
    map(~ {
      .x %>%
        compmus_normalise(method = norm) %>%
        enframe(name = "mfcc") %>%
        mutate(mfcc = mfcc - 1)
    }) %>%
    list_rbind(names_to = "time") %>%
    mutate(time = 1024 * time / 44100)
}

# Compute self-similarity matrices for chroma and timbre
compmus_self_similarity <- function(dat, feature, distance = "euclidean") {
  feature <- enquo(feature)
  compmus_long_distance(dat, dat, !!feature, distance)
}

# Function to compute pairwise distances
compmus_long_distance <- function(xdat, ydat, feature, distance = "euclidean") {
  feature <- enquo(feature)

  manhattan <- function(x, y) sum(abs(x - y))
  euclidean <- function(x, y) sqrt(sum((x - y)^2))
  chebyshev <- function(x, y) max(abs(x - y))
  pearson <- function(x, y) 1 - cor(x, y)
  cosine <- function(x, y) { 1 - sum(compmus_normalise(x, "euclidean") * compmus_normalise(y, "euclidean")) }
  angular <- function(x, y) 2 * acos(1 - cosine(x, y)) / pi
  aitchison <- function(x, y) { euclidean(compmus_normalise(x, "clr"), compmus_normalise(y, "clr")) }

  METHODS <- list(
    manhattan = manhattan, cityblock = manhattan, taxicab = manhattan, L1 = manhattan, totvar = manhattan,
    euclidean = euclidean, L2 = euclidean,
    chebyshev = chebyshev, maximum = chebyshev,
    pearson = pearson, correlation = pearson,
    cosine = cosine,
    angular = angular,
    aitchison = aitchison
  )

  if (!is.na(i <- pmatch(distance, names(METHODS)))) {
    dplyr::cross_join(
      xdat %>%
        rename(xtime = time) %>%
        tidyr::nest(.by = xtime) %>%
        transmute(
          xtime,
          x = map(data, ~ .x %>% arrange(!!feature) %>% pull(value))
        ) %>%
        arrange(xtime) %>%
        filter(row_number(xtime) %% 10 == 5),
      ydat %>%
        rename(ytime = time) %>%
        tidyr::nest(.by = ytime) %>%
        transmute(
          ytime,
          y = map(data, ~ .x %>% arrange(!!feature) %>% pull(value))
        ) %>%
        arrange(ytime) %>%
        filter(row_number(ytime) %% 10 == 5)
    ) %>%
      transmute(
        xtime,
        ytime,
        d = map2_dbl(x, y, METHODS[[i]])
      )
  } else {
    stop("The distance name is ambiguous or unsupported.")
  }
}

# Chroma and Timbre Matching Example
compmus_match_pitch_template <- function(dat, templates, method = "cosine", norm = "euclidean") {
  compmus_long_distance(
    dat %>%
      mutate(pitches = map(pitches, compmus_normalise, norm)),
    templates %>%
      transmute(
        start = name,
        duration = 1,
        pitches = map(template, compmus_normalise, norm)
      ),
    feature = pitches,
    method = method
  ) %>%
    transmute(
      start = xstart,
      duration = xduration,
      name = factor(ystart, levels = pluck(templates, "name")),
      d
    )
}

# Plot Self-Similarity Matrices
plot_similarity_matrix <- function(similarity_data, title) {
  ggplot(similarity_data, aes(x = xtime, y = ytime, fill = d)) +
    geom_tile() +
    scale_fill_gradient(low = "blue", high = "red") +
    labs(title = title, x = "Time", y = "Time", fill = "Distance") +
    theme_minimal()
}

# Example Execution: Load and Analyze a Track
file_path <- "/Users/thomas/Desktop/jsonTracks/thomas-m-1.json" # Update with actual file path

# Compute chroma and timbre
circshift <- function(vec, shift) {
  n <- length(vec)
  shift <- shift %% n  # Ensure the shift is within bounds
  c(tail(vec, -shift), head(vec, shift))
}

chroma_data <- compmus_chroma(file_path, norm = "euclidean")
mfcc_data <- compmus_mfccs(file_path, norm = "euclidean")

# Compute self-similarity matrices
chroma_similarity <- compmus_self_similarity(chroma_data, feature = pc, distance = "cosine")
mfcc_similarity <- compmus_self_similarity(mfcc_data, feature = mfcc, distance = "euclidean")

# Plot similarity matrices
plot_similarity_matrix(chroma_similarity, "Chroma Self-Similarity Matrix")
plot_similarity_matrix(mfcc_similarity, "Timbre (MFCC) Self-Similarity Matrix")


```

##  {data-width="600"}

### Other relevant information

Welcome to my **Portofolio** for the Computational Musicology course! The two tracks I have created were made using **flstudio**, about 2 years ago when I was getting started with music production.

```{r}

```

# Visualisation

## Column {data-width="650"}

### Chart A

```{r}

```

## Column {data-width="350"}

### Chart B

```{r}

```

### Chart C

```{r}

```

# Conclusion

## Column {data-width="650"}

### Chart A

```{r}

```

## Column {data-width="350"}

### Chart B

```{r}

```

### Chart C

```{r}

```
