---
title: "Portofolio for Computational Musicology 2025"
author: "Thomas Manu"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    orientation: columns
---

```{r setup, include=FALSE}
library(rjson)
library(tidyverse)
library(tuneR)
library(seewave)
library(ggplot2)
library(dplyr)
library(purrr)
library(jsonlite) 
```

# Introduction {.storyboard}

##  {data-width="400"}

### Welcome!

Welcome to my **Portofolio** for the Computational Musicology class! A little bit of a backstory on how I created the tracks. 

Track 1 was created on FlStudio, around 2 years ago when I was just starting music production. Growing up, the boom bap type of hiphop music was my gateway to other genres, and this track represents my teenage years. I used a sample taken from Splice, pitched it up and chopped it to create the sudden breaks in the vocals. The drum patterns were done in FlStudio. The BPM of the track is 90. 

Track 2 was also done in FlStudio. This 130 BPM track represents the Romanian minimal house genre, which I used to listen to a lot. 

##  {data-width="600"}

### Other relevant information


```{r}

```

# Visualisation {.storyboard}

##  {data-width="400"}

### Track 1: Chromagram 

```{r}
# Normalize vectors using various normalization strategies
compmus_normalise <- function(v, method = "euclidean") {
  harmonic <- function(v) v * sum(1 / (abs(v) + 1e-8))  # Prevent division by zero
  manhattan <- function(v) v / (sum(abs(v)) + 1e-8)
  euclidean <- function(v) v / sqrt(sum(v^2) + 1e-8)
  chebyshev <- function(v) v / (max(abs(v)) + 1e-8)
  clr <- function(v) { lv <- log(v + 1e-8); lv - mean(lv) }
  softmax <- function(v) exp(v) / sum(exp(v))
  
  METHODS <- list(
    identity = identity,
    harmonic = harmonic,
    manhattan = manhattan, L1 = manhattan,
    euclidean = euclidean, L2 = euclidean,
    chebyshev = chebyshev, maximum = chebyshev,
    aitchison = clr, clr = clr,
    softmax = softmax
  )

  i <- pmatch(method, names(METHODS))
  if (!is.na(i)) {
    METHODS[[names(METHODS)[i]]](v)
  } else {
    stop("The method name is ambiguous or unsupported.")
  }
}

# Function to perform circular shift
circshift <- function(vec, shift) {
  n <- length(vec)
  shift <- shift %% n  # Ensure the shift is within bounds
  if (shift == 0) return(vec)
  c(tail(vec, n - shift), head(vec, shift))
}

# Compute chroma features
compmus_chroma <- function(file, norm = "identity") {
  if (!file.exists(file)) stop("File does not exist: ", file)
  
  # Corrected way to load JSON data
  json_data <- fromJSON(readLines(file), flatten = TRUE)
  
  # Check if JSON contains the expected structure
  if (!("tonal" %in% names(json_data)) || !("hpcp" %in% names(json_data$tonal))) {
    stop("JSON structure is invalid: missing 'tonal$hpcp'")
  }

  json_data$tonal$hpcp %>%
    map(~ .x %>%
           circshift(-8) %>%
           matrix(nrow = 3, ncol = 12, byrow = TRUE) %>%
           colMeans() %>%
           compmus_normalise(method = norm) %>%
           enframe(name = "pc") %>%
           mutate(pc = as.numeric(pc) - 1)
    ) %>%
    list_rbind(names_to = "time") %>%
    mutate(time = 2048 * as.numeric(time) / 44100)
}

# Example Execution: Load and Analyze a Track
file_path <- "/Users/thomas/Desktop/jsonTracks/thomas-m-1.json" # Update with actual file path

# Compute chroma data
chroma_data <- compmus_chroma(file_path, norm = "euclidean")

# Plot chromagram
ggplot(chroma_data, aes(x = time, y = pc, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "red", high = "blue") +
  labs(title = "Chromagram", x = "Time (s)", y = "Pitch Class", fill = "Magnitude") +
  theme_minimal()
```

### Track 1: Timbre Self-Similarity Matrix

```{r}
# Compute MFCCs (Timbre)
compmus_mfccs <- function(file, norm = "identity") {
  if (!file.exists(file)) stop("File does not exist: ", file)
  
  # Corrected JSON loading
  json_data <- fromJSON(readLines(file), flatten = TRUE)
  
  # Ensure JSON contains expected structure
  if (!("lowlevel" %in% names(json_data)) || !("mfcc" %in% names(json_data$lowlevel))) {
    stop("JSON structure is invalid: missing 'lowlevel$mfcc'")
  }

  json_data$lowlevel$mfcc %>%
    map(~ .x %>%
           compmus_normalise(method = norm) %>%
           enframe(name = "mfcc") %>%
           mutate(mfcc = as.numeric(mfcc) - 1)
    ) %>%
    list_rbind(names_to = "time") %>%
    mutate(time = 1024 * as.numeric(time) / 44100)
}

# Compute pairwise distances
compmus_long_distance <- function(xdat, ydat, feature, distance = "euclidean") {
  feature <- enquo(feature)

  manhattan <- function(x, y) sum(abs(x - y))
  euclidean <- function(x, y) sqrt(sum((x - y)^2))
  chebyshev <- function(x, y) max(abs(x - y))
  pearson <- function(x, y) 1 - cor(x, y)
  cosine <- function(x, y) { 1 - sum(compmus_normalise(x, "euclidean") * compmus_normalise(y, "euclidean")) }
  angular <- function(x, y) 2 * acos(1 - cosine(x, y)) / pi
  aitchison <- function(x, y) { euclidean(compmus_normalise(x, "clr"), compmus_normalise(y, "clr")) }

  METHODS <- list(
    manhattan = manhattan, cityblock = manhattan, taxicab = manhattan, L1 = manhattan, totvar = manhattan,
    euclidean = euclidean, L2 = euclidean,
    chebyshev = chebyshev, maximum = chebyshev,
    pearson = pearson, correlation = pearson,
    cosine = cosine,
    angular = angular,
    aitchison = aitchison
  )

  i <- pmatch(distance, names(METHODS))
  if (!is.na(i)) {
    dplyr::cross_join(
      xdat %>%
        rename(xtime = time) %>%
        tidyr::nest(.by = xtime) %>%
        transmute(
          xtime,
          x = map(data, ~ .x %>% arrange(!!feature) %>% pull(value))
        ) %>%
        arrange(xtime) %>%
        filter(row_number(xtime) %% 10 == 5),
      ydat %>%
        rename(ytime = time) %>%
        tidyr::nest(.by = ytime) %>%
        transmute(
          ytime,
          y = map(data, ~ .x %>% arrange(!!feature) %>% pull(value))
        ) %>%
        arrange(ytime) %>%
        filter(row_number(ytime) %% 10 == 5)
    ) %>%
      transmute(
        xtime,
        ytime,
        d = map2_dbl(x, y, METHODS[[i]])
      )
  } else {
    stop("The distance name is ambiguous or unsupported.")
  }
}

# Compute self-similarity matrices
compmus_self_similarity <- function(dat, feature, distance = "euclidean") {
  feature <- enquo(feature)
  compmus_long_distance(dat, dat, !!feature, distance)
}

# Example Execution: Load and Analyze a Track
file_path <- "/Users/thomas/Desktop/jsonTracks/thomas-m-1.json" # Update with actual file path

# Compute MFCC data
mfcc_data <- compmus_mfccs(file_path, norm = "euclidean")

# Compute self-similarity matrix
mfcc_similarity <- compmus_self_similarity(mfcc_data, feature = mfcc, distance = "euclidean")

# Plot similarity matrix
ggplot(mfcc_similarity, aes(x = xtime, y = ytime, fill = d)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "yellow") +
  labs(title = "Timbre Self-Similarity Matrix", x = "Time", y = "Time", fill = "Distance") +
  theme_minimal()

``` 

### Track 2: Chromagram 
```{r}
# Normalize vectors using various normalization strategies
compmus_normalise <- function(v, method = "euclidean") {
  harmonic <- function(v) v * sum(1 / (abs(v) + 1e-8))  # Prevent division by zero
  manhattan <- function(v) v / (sum(abs(v)) + 1e-8)
  euclidean <- function(v) v / sqrt(sum(v^2) + 1e-8)
  chebyshev <- function(v) v / (max(abs(v)) + 1e-8)
  clr <- function(v) { lv <- log(v + 1e-8); lv - mean(lv) }
  softmax <- function(v) exp(v) / sum(exp(v))
  
  METHODS <- list(
    identity = identity,
    harmonic = harmonic,
    manhattan = manhattan, L1 = manhattan,
    euclidean = euclidean, L2 = euclidean,
    chebyshev = chebyshev, maximum = chebyshev,
    aitchison = clr, clr = clr,
    softmax = softmax
  )

  i <- pmatch(method, names(METHODS))
  if (!is.na(i)) {
    METHODS[[names(METHODS)[i]]](v)
  } else {
    stop("The method name is ambiguous or unsupported.")
  }
}

# Function to perform circular shift
circshift <- function(vec, shift) {
  n <- length(vec)
  shift <- shift %% n  # Ensure the shift is within bounds
  if (shift == 0) return(vec)
  c(tail(vec, n - shift), head(vec, shift))
}

# Compute chroma features
compmus_chroma <- function(file, norm = "identity") {
  if (!file.exists(file)) stop("File does not exist: ", file)

  # Correct JSON loading
  json_data <- fromJSON(readLines(file), flatten = TRUE)

  # Ensure JSON structure is valid
  if (!("tonal" %in% names(json_data)) || !("hpcp" %in% names(json_data$tonal))) {
    stop("JSON structure is invalid: missing 'tonal$hpcp'")
  }

  json_data$tonal$hpcp %>%
    map(~ .x %>%
           circshift(-8) %>%
           matrix(nrow = 3, ncol = 12, byrow = TRUE) %>%
           colMeans() %>%
           compmus_normalise(method = norm) %>%
           enframe(name = "pc") %>%
           mutate(pc = as.numeric(pc) - 1)
    ) %>%
    list_rbind(names_to = "time") %>%
    mutate(time = 2048 * as.numeric(time) / 44100)
}

# Example Execution: Load and Analyze a Track
file_path <- "/Users/thomas/Desktop/jsonTracks/thomas-m-2.json" # Update with actual file path

# Compute chroma data
chroma_data <- compmus_chroma(file_path, norm = "euclidean")

# Plot chromagram
ggplot(chroma_data, aes(x = time, y = pc, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(title = "Chromagram", x = "Time (s)", y = "Pitch Class", fill = "Magnitude") +
  theme_minimal()
``` 


### Track 2: Timbre Self-Similarity Matrix

```{r} 
# Compute MFCCs (Timbre)
compmus_mfccs <- function(file, norm = "identity") {
  if (!file.exists(file)) stop("File does not exist: ", file)
  
  # Correct JSON loading
  json_data <- fromJSON(readLines(file), flatten = TRUE)
  
  # Ensure JSON contains expected structure
  if (!("lowlevel" %in% names(json_data)) || !("mfcc" %in% names(json_data$lowlevel))) {
    stop("JSON structure is invalid: missing 'lowlevel$mfcc'")
  }

  json_data$lowlevel$mfcc %>%
    map(~ .x %>%
           compmus_normalise(method = norm) %>%
           enframe(name = "mfcc") %>%
           mutate(mfcc = as.numeric(mfcc) - 1)
    ) %>%
    list_rbind(names_to = "time") %>%
    mutate(time = 1024 * as.numeric(time) / 44100)
}

# Compute pairwise distances
compmus_long_distance <- function(xdat, ydat, feature, distance = "euclidean") {
  feature <- enquo(feature)

  manhattan <- function(x, y) sum(abs(x - y))
  euclidean <- function(x, y) sqrt(sum((x - y)^2))
  chebyshev <- function(x, y) max(abs(x - y))
  pearson <- function(x, y) 1 - cor(x, y)
  cosine <- function(x, y) { 1 - sum(compmus_normalise(x, "euclidean") * compmus_normalise(y, "euclidean")) }
  angular <- function(x, y) 2 * acos(1 - cosine(x, y)) / pi
  aitchison <- function(x, y) { euclidean(compmus_normalise(x, "clr"), compmus_normalise(y, "clr")) }

  METHODS <- list(
    manhattan = manhattan, cityblock = manhattan, taxicab = manhattan, L1 = manhattan, totvar = manhattan,
    euclidean = euclidean, L2 = euclidean,
    chebyshev = chebyshev, maximum = chebyshev,
    pearson = pearson, correlation = pearson,
    cosine = cosine,
    angular = angular,
    aitchison = aitchison
  )

  i <- pmatch(distance, names(METHODS))
  if (!is.na(i)) {
    dplyr::cross_join(
      xdat %>%
        rename(xtime = time) %>%
        tidyr::nest(.by = xtime) %>%
        transmute(
          xtime,
          x = map(data, ~ .x %>% arrange(!!feature) %>% pull(value))
        ) %>%
        arrange(xtime) %>%
        filter(row_number(xtime) %% 10 == 5),
      ydat %>%
        rename(ytime = time) %>%
        tidyr::nest(.by = ytime) %>%
        transmute(
          ytime,
          y = map(data, ~ .x %>% arrange(!!feature) %>% pull(value))
        ) %>%
        arrange(ytime) %>%
        filter(row_number(ytime) %% 10 == 5)
    ) %>%
      transmute(
        xtime,
        ytime,
        d = map2_dbl(x, y, METHODS[[names(METHODS)[i]]])
      )
  } else {
    stop("The distance name is ambiguous or unsupported.")
  }
}

# Compute self-similarity matrices
compmus_self_similarity <- function(dat, feature, distance = "euclidean") {
  feature <- enquo(feature)
  compmus_long_distance(dat, dat, !!feature, distance)
}

# Example Execution: Load and Analyze a Track
file_path <- "/Users/thomas/Desktop/jsonTracks/thomas-m-2.json" # Update with actual file path

# Compute MFCC data
mfcc_data <- compmus_mfccs(file_path, norm = "euclidean")

# Compute self-similarity matrix
mfcc_similarity <- compmus_self_similarity(mfcc_data, feature = mfcc, distance = "euclidean")

# Plot similarity matrix
ggplot(mfcc_similarity, aes(x = xtime, y = ytime, fill = d)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "yellow") +
  labs(title = "Timbre Self-Similarity Matrix", x = "Time", y = "Time", fill = "Distance") +
  theme_minimal()

```

### Track 1: Analysis of chords 

```{r}
# Function to compute chord analysis from JSON files
compmus_chords <- function(file) {
  json_data <- fromJSON("/Users/thomas/Desktop/jsonTracks/thomas-m-1.json")
  
  # Extract harmonic pitch class profile (HPCP) data
  chord_data <- json_data$tonal$chords_key %>%
    enframe(name = "time", value = "key") %>%
    mutate(
      time = 2048 * (time - 1) / 44100,  # Convert frame index to seconds
      chords_scale = json_data$tonal$chords_scale[time + 1],
      chords_progression = paste(key, chords_scale, sep = " ")
    )
  
  return(chord_data)
}

# ---- Track 1 Analysis ----

# Load and Analyze Chords from Track 1
file_path_1 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-1.json" # Update with actual file path
chord_data_1 <- compmus_chords(file_path_1)

# Plot chord progression for Track 1
ggplot(chord_data_1, aes(x = time, y = chords_progression, fill = chords_progression)) +
  geom_tile() +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title = "Chord Analysis - Track 1",
    x = "Time (s)",
    y = "Chord Progression",
    fill = "Chords"
  ) +
  theme_minimal()
```

### Track 2: Analysis of chords 

```{r}
# Function to compute chord analysis from JSON files
compmus_chords <- function(file) {
  json_data <- fromJSON("/Users/thomas/Desktop/jsonTracks/thomas-m-2.json")
  
  # Extract harmonic pitch class profile (HPCP) data
  chord_data <- json_data$tonal$chords_key %>%
    enframe(name = "time", value = "key") %>%
    mutate(
      time = 2048 * (time - 1) / 44100,  # Convert frame index to seconds
      chords_scale = json_data$tonal$chords_scale[time + 1],
      chords_progression = paste(key, chords_scale, sep = " ")
    )
  
  return(chord_data) 
}
# ---- Track 2 Analysis ----

# Load and Analyze Chords from Track 2
file_path_2 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-2.json" # Update with actual file path
chord_data_2 <- compmus_chords(file_path_2)

# Plot chord progression for Track 2
ggplot(chord_data_2, aes(x = time, y = chords_progression, fill = chords_progression)) +
  geom_tile() +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title = "Chord Analysis - Track 2",
    x = "Time (s)",
    y = "Chord Progression",
    fill = "Chords"
  ) +
  theme_minimal()
```

# Conclusion

## Column {data-width="650"}

### Chart A

```{r}

```

## Column {data-width="350"}

### Chart B

```{r}

```

### Chart C

```{r}

```
