---
title: "Portofolio for Computational Musicology 2025"
author: "Thomas Manu"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    orientation: columns
---

```{r setup, include=FALSE}
library(rjson)
library(tidyverse)
library(tuneR)
library(seewave)
library(ggplot2)
library(dplyr)
library(purrr)
library(jsonlite) 
```

# Introduction {.storyboard}

##  {data-width="400"}

### Welcome!

Welcome to my **Portofolio** for the Computational Musicology class! A little bit of a backstory on how I created the tracks. 

Track 1 was created on FlStudio, around 2 years ago when I was just starting music production. Growing up, the boom bap type of hiphop music was my gateway to other genres, and this track represents my teenage years. I used a sample taken from Splice, pitched it up and chopped it to create the sudden breaks in the vocals. The drum patterns were done in FlStudio. The BPM of the track is 90. 

Track 2 was also done in FlStudio. This 130 BPM track represents the Romanian minimal house genre, which I used to listen to a lot.

**My computer couldn't handle rendering all the chromagrams and SSMs, so I included only  HW 11.** 

##  {data-width="600"}

### Other relevant information

Global Functions 
```{r}
# Function to normalize vectors
compmus_normalise <- function(v, method = "euclidean") {
  METHODS <- list(
    identity = identity,
    harmonic = function(v) v * sum(1 / abs(v)),
    manhattan = function(v) v / sum(abs(v)),
    euclidean = function(v) v / sqrt(sum(v^2)),
    chebyshev = function(v) v / max(abs(v)),
    clr = function(v) { lv <- log(v); lv - mean(lv) },
    softmax = function(v) exp(v) / sum(exp(v))
  )
  
  if (method %in% names(METHODS)) {
    return(METHODS[[method]](v))
  } else {
    stop("The method name is ambiguous or unsupported.")
  }
}

# Function to compute chroma features
compmus_chroma <- function(file, norm = "identity") {
  json_text <- paste(readLines(file, warn = FALSE), collapse = "")
  json_data <- fromJSON(txt = json_text)

  if (is.null(json_data$tonal$hpcp) || length(json_data$tonal$hpcp) == 0) {
    stop("Error: No valid chroma data found in JSON file!")
  }
  
  chroma_matrix <- json_data$tonal$hpcp  # Extract raw chroma data

  # Ensure chroma_matrix is a valid list of vectors
  if (is.list(chroma_matrix) && all(sapply(chroma_matrix, is.numeric))) {
    chroma_matrix <- do.call(rbind, chroma_matrix)  # Convert list of vectors to matrix
  } else {
    stop("Error: Chroma data format is incorrect!")
  }
  
  # Ensure we have the expected number of pitch classes (12)
  if (ncol(chroma_matrix) != 12) {
    stop("Error: Chroma matrix does not have 12 pitch classes!")
  }

  # Normalize and format data
  chroma_data <- as_tibble(chroma_matrix) %>%
    mutate(time = seq(0, by = (2048 / 44100), length.out = nrow(chroma_matrix))) %>%
    pivot_longer(cols = starts_with("V"), names_to = "pc", values_to = "value") %>%
    mutate(pc = as.integer(gsub("V", "", pc)) - 1)  # Convert column names to pitch class numbers
  
  return(chroma_data)
}

# Function to compute Chordograms & Keygrams
compmus_extract <- function(file, feature) {
  json_text <- paste(readLines(file, warn = FALSE), collapse = "")
  json_data <- fromJSON(txt = json_text)
  
  if (feature == "chords") {
    key_names <- unlist(json_data$tonal$chords_key, use.names = FALSE)
    key_confidence <- unlist(json_data$tonal$chords_strength, use.names = FALSE)
  } else if (feature == "key") {
    key_names <- unlist(json_data$tonal$key_krumhansl$key, use.names = FALSE)
    key_confidence <- unlist(json_data$tonal$key_krumhansl$strength, use.names = FALSE)
  } else {
    stop("Invalid feature type. Choose 'chords' or 'key'.")
  }
  
  num_keys <- length(key_names)
  time_values <- seq(0, json_data$metadata$audio_properties$length, length.out = num_keys)
  
  return(tibble(Time = time_values, Feature = key_names, Intensity = key_confidence))
}

# Define file paths for both tracks
file_path_1 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-1.json"
file_path_2 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-2.json"
``` 

### HW 11: Histogram of Tempi
```{r}
corpus_files <- list.files("/Users/thomas/Desktop/jsonTracks/", full.names = TRUE, pattern = "*.json")
corpus_tempi <- map_dbl(corpus_files, function(file) {
  json_text <- paste(readLines(file, warn = FALSE), collapse = "")
  json_data <- fromJSON(txt = json_text)
  if (!is.null(json_data$rhythm$bpm)) {
    return(json_data$rhythm$bpm)
  } else {
    return(NA)  
  }
})

corpus_tempi <- corpus_tempi[!is.na(corpus_tempi)]

ggplot(tibble(BPM = corpus_tempi), aes(x = BPM)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "white", alpha = 0.7) +
  labs(title = "Histogram of Tempi in the Class Corpus", x = "BPM", y = "Frequency") +
  theme_minimal()
```

### Tempograms for track 1
```{r}
tempogram_data_1 <- compmus_extract(file_path_1, "key")

ggplot(tempogram_data_1, aes(x = Time, y = Intensity)) +
  geom_line(color = "blue") +
  labs(title = "Tempogram - Track 1", x = "Time (s)", y = "BPM Confidence") +
  theme_minimal()
``` 

### Tempograms for track 2
```{r}
tempogram_data_2 <- compmus_extract(file_path_2, "key")

ggplot(tempogram_data_2, aes(x = Time, y = Intensity)) +
  geom_line(color = "blue") +
  labs(title = "Tempogram - Track 2", x = "Time (s)", y = "BPM Confidence") +
  theme_minimal()
``` 


# Visualisation {.storyboard}

##  {data-width="400"}

### Track 1: Chromagram 

```{r}

```

### Track 1: Timbre Self-Similarity Matrix

```{r}

``` 

### Track 2: Chromagram 
```{r}

``` 

### Track 2: Timbre Self-Similarity Matrix

```{r} 

```

### Track 1: Chordogram 

```{r}

```

### Track 2: Chordogram

```{r}

```

### Histogram of Tempi
```{r}

```

### Tempograms for track 1
```{r}

``` 

### Tempograms for track 2
```{r}

``` 

# Conclusion

## Column {data-width="650"}

### Chart A

```{r}

```

## Column {data-width="350"}

### Chart B

```{r}

```

### Chart C

```{r}

```
