---
title: "Portofolio for Computational Musicology 2025"
author: "Thomas Manu"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    orientation: columns
---

```{r setup, include=FALSE}
library(rjson)
library(tidyverse)
library(tuneR)
library(seewave)
library(ggplot2)
library(dplyr)
library(purrr)
library(jsonlite) 
```

# Introduction {.storyboard}

##  {data-width="400"}

### Welcome!

Welcome to my **Portofolio** for the Computational Musicology class! A little bit of a backstory on how I created the tracks. 

Track 1 was created on FlStudio, around 2 years ago when I was just starting music production. Growing up, the boom bap type of hiphop music was my gateway to other genres, and this track represents my teenage years. I used a sample taken from Splice, pitched it up and chopped it to create the sudden breaks in the vocals. The drum patterns were done in FlStudio. The BPM of the track is 90. 

Track 2 was also done in FlStudio. This 130 BPM track represents the Romanian minimal house genre, which I used to listen to a lot.

**My computer couldn't handle rendering all the chromagrams and SSMs, so I included only  the Chord Analysis. Please scroll to the last 2 sections on the Visualization page.** 

##  {data-width="600"}

### Other relevant information


```{r}

```

# Visualisation {.storyboard}

##  {data-width="400"}

### Track 1: Chromagram 


```{r}

```

### Track 1: Timbre Self-Similarity Matrix

```{r}

``` 

### Track 2: Chromagram 
```{r}

``` 


### Track 2: Timbre Self-Similarity Matrix

```{r} 

```

### Track 1: Analysis of chords 

```{r}
# Function to compute chord analysis from JSON files
compmus_chords <- function(file) {
  json_data <- fromJSON("/Users/thomas/Desktop/jsonTracks/thomas-m-1.json")
  
  # Extract harmonic pitch class profile (HPCP) data
  chord_data <- json_data$tonal$chords_key %>%
    enframe(name = "time", value = "key") %>%
    mutate(
      time = 2048 * (time - 1) / 44100,  # Convert frame index to seconds
      chords_scale = json_data$tonal$chords_scale[time + 1],
      chords_progression = paste(key, chords_scale, sep = " ")
    )
  
  return(chord_data) 
}

# ---- Track 1 Analysis ----

# Load and Analyze Chords from Track 1
file_path_1 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-1.json" # Update with actual file path
chord_data_1 <- compmus_chords(file_path_1)

# Plot chord progression for Track 1
ggplot(chord_data_1, aes(x = time, y = chords_progression, fill = chords_progression)) +
  geom_tile() +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title = "Chord Analysis - Track 1",
    x = "Time (s)",
    y = "Chord Progression",
    fill = "Chords"
  ) +
  theme_minimal()
```

### Track 2: Analysis of chords 

```{r}
# Function to compute chord analysis from JSON files
compmus_chords <- function(file) {
  json_data <- fromJSON("/Users/thomas/Desktop/jsonTracks/thomas-m-2.json")
  
  # Extract harmonic pitch class profile (HPCP) data
  chord_data <- json_data$tonal$chords_key %>%
    enframe(name = "time", value = "key") %>%
    mutate(
      time = 2048 * (time - 1) / 44100,  # Convert frame index to seconds
      chords_scale = json_data$tonal$chords_scale[time + 1],
      chords_progression = paste(key, chords_scale, sep = " ")
    )
  
  return(chord_data) 
}
# ---- Track 2 Analysis ----

# Load and Analyze Chords from Track 2
file_path_2 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-2.json" # Update with actual file path
chord_data_2 <- compmus_chords(file_path_2)

# Plot chord progression for Track 2
ggplot(chord_data_2, aes(x = time, y = chords_progression, fill = chords_progression)) +
  geom_tile() +
  scale_fill_brewer(palette = "Set3") +
  labs(
    title = "Chord Analysis - Track 2",
    x = "Time (s)",
    y = "Chord Progression",
    fill = "Chords"
  ) +
  theme_minimal() 
```

# Conclusion

## Column {data-width="650"}

### Chart A

```{r}

```

## Column {data-width="350"}

### Chart B

```{r}

```

### Chart C

```{r}

```
