---
title: "Portofolio for Computational Musicology 2025"
author: "Thomas Manu"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    orientation: columns
---

```{r setup, include=FALSE}
library(rjson)
library(tidyverse)
library(tuneR)
library(seewave)
library(ggplot2)
library(dplyr)
library(purrr)
library(jsonlite) 
library(tibble)
library(stringr)
```

# Introduction {.storyboard}

##  {data-width="400"}

### Welcome!

Welcome to my **Portofolio** for the Computational Musicology class! A little bit of a backstory on how I created the tracks. 

Track 1 was created on FlStudio, around 2 years ago when I was just starting music production. Growing up, the boom bap type of hiphop music was my gateway to other genres, and this track represents my teenage years. I used a sample taken from Splice, pitched it up and chopped it to create the sudden breaks in the vocals. The drum patterns were done in FlStudio. The BPM of the track is 90. 

Track 2 was also done in FlStudio. This 130 BPM track represents the Romanian minimal house genre, which I used to listen to a lot growing up. 

I hope you'll enjoy this computational journey! 

##  {data-width="600"}

### Other relevant information

Global Functions 
```{r}
# Function to normalize vectors
compmus_normalise <- function(v, method = "euclidean") {
  METHODS <- list(
    identity = identity,
    harmonic = function(v) v * sum(1 / abs(v)),
    manhattan = function(v) v / sum(abs(v)),
    euclidean = function(v) v / sqrt(sum(v^2)),
    chebyshev = function(v) v / max(abs(v)),
    clr = function(v) { lv <- log(v); lv - mean(lv) },
    softmax = function(v) exp(v) / sum(exp(v))
  )
  
  if (method %in% names(METHODS)) {
    return(METHODS[[method]](v))
  } else {
    stop("The method name is ambiguous or unsupported.")
  }
}

# Function to compute chroma features
compmus_chroma <- function(file, norm = "identity") {
  json_text <- paste(readLines(file, warn = FALSE), collapse = "")
  json_data <- fromJSON(txt = json_text)

  if (is.null(json_data$tonal$hpcp) || length(json_data$tonal$hpcp) == 0) {
    stop("Error: No valid chroma data found in JSON file!")
  }
  
  chroma_matrix <- json_data$tonal$hpcp  # Extract raw chroma data

  # Ensure chroma_matrix is a valid list of vectors
  if (is.list(chroma_matrix) && all(sapply(chroma_matrix, is.numeric))) {
    chroma_matrix <- do.call(rbind, chroma_matrix)  # Convert list of vectors to matrix
  } else {
    stop("Error: Chroma data format is incorrect!")
  }
  
  # Ensure we have the expected number of pitch classes (12)
  if (ncol(chroma_matrix) != 12) {
    stop("Error: Chroma matrix does not have 12 pitch classes!")
  }

  # Normalize and format data
  chroma_data <- as_tibble(chroma_matrix) %>%
    mutate(time = seq(0, by = (2048 / 44100), length.out = nrow(chroma_matrix))) %>%
    pivot_longer(cols = starts_with("V"), names_to = "pc", values_to = "value") %>%
    mutate(pc = as.integer(gsub("V", "", pc)) - 1)  # Convert column names to pitch class numbers
  
  return(chroma_data)
}

# Function to compute Chordograms & Keygrams
compmus_extract <- function(file, feature) {
  json_text <- paste(readLines(file, warn = FALSE), collapse = "")
  json_data <- fromJSON(txt = json_text)
  
  if (feature == "chords") {
    key_names <- unlist(json_data$tonal$chords_key, use.names = FALSE)
    key_confidence <- unlist(json_data$tonal$chords_strength, use.names = FALSE)
  } else if (feature == "key") {
    key_names <- unlist(json_data$tonal$key_krumhansl$key, use.names = FALSE)
    key_confidence <- unlist(json_data$tonal$key_krumhansl$strength, use.names = FALSE)
  } else {
    stop("Invalid feature type. Choose 'chords' or 'key'.")
  }
  
  num_keys <- length(key_names)
  time_values <- seq(0, json_data$metadata$audio_properties$length, length.out = num_keys)
  
  return(tibble(Time = time_values, Feature = key_names, Intensity = key_confidence))
}

# Define file paths for both tracks
file_path_1 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-1.json"
file_path_2 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-2.json"
``` 

# Visualisation {.storyboard}

##  {data-width="400"}

### Track 1: Chromagram 

```{r}
# === Function: Compute chromagram from HPCP ===
compmus_chroma <- function(file, norm = "euclidean") {
  # Read file
  json_text <- paste(readLines(file, warn = FALSE), collapse = "")
  json_data <- fromJSON(txt = json_text)

  # Extract HPCP matrix
  hpcp <- json_data$tonal$hpcp

  # Convert to tibble with time and pitch class (pc)
  chroma_data <- hpcp %>%
    map_dfr(~ enframe(.x, name = "pc", value = "value"), .id = "frame") %>%
    mutate(
      pc = as.numeric(pc) - 1,
      frame = as.numeric(frame),
      time = frame * (2048 / 44100)
    )

  return(chroma_data)
}

# === Set filepath ===
file_path_1 <- "/Users/thomas/Desktop/jsonTracks/thomas-m-1.json"

# === Run ===
chroma_data_1 <- compmus_chroma(file_path_1)

# === Plot ===
ggplot(chroma_data_1, aes(x = time, y = pc, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "darkred", high = "lightblue") +
  labs(
    title = "Chromagram – Track 1",
    x = "Time (s)", y = "Pitch Class", fill = "Magnitude"
  ) +
  theme_minimal()
```

### Track 1: MFCC-Based Cepstrogram

```{r}
# --- MFCC Extraction ---
compmus_mfccs <- function(file, norm = "identity") {
  json_text <- paste(readLines(file, warn = FALSE), collapse = "")
  json_data <- fromJSON(txt = json_text)

  mfcc_matrix <- json_data$lowlevel$mfcc

  mfcc_data <- mfcc_matrix %>%
    map(~ compmus_normalise(.x, method = norm)) %>%
    map_dfr(~ enframe(.x, name = "mfcc", value = "value"), .id = "time") %>%
    mutate(
      time = as.numeric(time) * (1024 / 44100),
      mfcc = as.numeric(mfcc) - 1
    )

  return(mfcc_data)
}

# --- Apply to Track 1 ---
mfcc_data_1 <- compmus_mfccs("/Users/thomas/Desktop/jsonTracks/thomas-m-1.json", norm = "euclidean")

# --- Plot Cepstrogram ---
ggplot(mfcc_data_1, aes(x = time, y = mfcc, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "gold") +
  labs(title = "Cepstrogram (MFCCs) – Track 1", x = "Time (s)", y = "MFCC", fill = "Value") +
  theme_minimal()

``` 

### Track 2: Chromagram 
```{r}
# Reuse chroma function from earlier
chroma_data_2 <- compmus_chroma(file_path_2)

# Plot chromagram
ggplot(chroma_data_2, aes(x = time, y = pc, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "darkred", high = "lightblue") +
  labs(
    title = "Chromagram – Track 2",
    x = "Time (s)", y = "Pitch Class", fill = "Magnitude"
  ) +
  theme_minimal()

``` 

### Track 2: MFCC-Based Cepstrogram
```{r}
# Extract MFCCs
mfcc_data_2 <- compmus_mfccs(file_path_2, norm = "euclidean")

# Plot
ggplot(mfcc_data_2, aes(x = time, y = mfcc, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "gold") +
  labs(
    title = "Cepstrogram (MFCCs) – Track 2",
    x = "Time (s)", y = "MFCC", fill = "Value"
  ) +
  theme_minimal()

```

### Track 1: Chordogram 

```{r}
# Extract chord data for Track 1
chords_1 <- compmus_extract(file_path_1, feature = "chords")

# Plot chordogram
ggplot(chords_1, aes(x = Time, y = Feature, fill = Intensity)) +
  geom_tile() +
  scale_fill_gradient(low = "darkred", high = "lightblue") +
  labs(
    title = "Chordogram – Track 1",
    x = "Time (s)",
    y = "Chord",
    fill = "Confidence"
  ) +
  theme_minimal()

```

### Track 2: Chordogram

```{r}
# Extract chord data for Track 2
chords_2 <- compmus_extract(file_path_2, feature = "chords")

# Plot chordogram
ggplot(chords_2, aes(x = Time, y = Feature, fill = Intensity)) +
  geom_tile() +
  scale_fill_gradient(low = "darkred", high = "lightblue") +
  labs(
    title = "Chordogram – Track 2",
    x = "Time (s)",
    y = "Chord",
    fill = "Confidence"
  ) +
  theme_minimal()

```


### Track 1: Keygram 
```{r}
# Extract key data for Track 1
keys_1 <- compmus_extract(file_path_1, feature = "key")

# Plot keygram
ggplot(keys_1, aes(x = Time, y = Feature, fill = Intensity)) +
  geom_tile() +
  scale_fill_gradient(low = "darkblue", high = "gold") +
  labs(
    title = "Keygram – Track 1",
    x = "Time (s)",
    y = "Key",
    fill = "Confidence"
  ) +
  theme_minimal()

```

### Track 2: Keygram 
```{r}
# Extract key data for Track 2
keys_2 <- compmus_extract(file_path_2, feature = "key")

# Plot keygram
ggplot(keys_2, aes(x = Time, y = Feature, fill = Intensity)) +
  geom_tile() +
  scale_fill_gradient(low = "darkblue", high = "gold") +
  labs(
    title = "Keygram – Track 2",
    x = "Time (s)",
    y = "Key",
    fill = "Confidence"
  ) +
  theme_minimal()

```

### Tempogram for my 2 tracks

No tempo (BPM) data was available in the JSON output for either of my tracks. Therefore, tempogram analysis could not be performed.

# Conclusion

## Column {data-width="650"}

### Chart A

```{r}

```

## Column {data-width="350"}

### Chart B

```{r}

```

### Chart C

```{r}

```
